PhD Title \\ PhD Thesis

### 2024-06-07 Latex_files/Thesis/chapters/chapter 2/main.tex
updating main and chapters using
reformatting chapter2

\chapter{Theory and methods}
The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the work being done with 
optical tweezers in particle classification, and how tweezing has been 
used to investigate nucleation events.

\section{Working Principle}
Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects. If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself. While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory. Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}. This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0. However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.

@@

\section{Electromagnetism and optical tweezing}
Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
	&\nabla \cdot \mathbf{E} = \frac{\rho_ v}{\epsilon_0} \\
	&\nabla \cdot \mathbf{B} = 0 \\
	&\nabla \times \mathbf{E}  = \frac{\delta \mathbf{B}}{\delta t} \\
	&\nabla \times \mathbf{B} = j_0 +\frac{\delta \mathbf{E}}{\delta t}    
\end{align}
These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

@@

\subsection{Ray-Optics Regime}
The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law. Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam. Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge. As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:

\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}

Where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference in refractive indices between the solution and the target particle, and $P_i$ is the power of the individual ray. For a beam with a Gaussian intensity distribution $P_i$ will fall off as you move from the centre of the beam. The ray optics model is ideal for

\subsection{Lornez-Mie Theory}
The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{fixed_polarisation.png}
\end{figure}

>>>>

\chapter{Theory and methods}

The two key areas of this project cover two widely different fields; 
crystallisation theory, which is covered in the previous introduction 
section, and tweezing theory. The following chapter summarises the 
working principles behind optical tweezing, the work being done with 
optical tweezers in particle classification, and how tweezing has been 
used to investigate nucleation events.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Working Principle}

Optical tweezing operates on the principle that light carries both linear and angular momentum, which can be transferred to other objects. If light is reflected or refracted by a medium part of the light's momentum is transferred to the medium itself. While at large scales this effect is insignificant in most cases, at micron scales the forces imparted by individual photons begin to become significant enough to influence its trajectory. Ashkin's work into optical tweezing found that micron sized entities could be spatially locked by aiming a laser with a Gaussian intensity profile directly upward at a the cell while being suspended in a liquid medium \cite{Ashkin1970}. This is demonstrated diagram below where we have a simple sphere trapped by a Gaussian beam, if we consider two rays of light hitting our sphere we see that while in the centre each ray is refracted in an equal opposite direction to one another and so the net force imparted is 0. However when displaced the net force is unbalanced and points towards the centre of the trap; this principle can be generalised to 3 degrees of freedom if we consider a focused beam instead of a para-axial beam, typically the trap will be weaker along the beam axis compared to the transverse trap strength.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Electromagnetism and optical tweezing}

Proper understanding of optical tweezing requires an understanding of how the trapped particle interacts with the trapping laser. From a electromagnetism perspective the laser creates a spatially and temporally coherent electric field that scatters light off of a trapped particle. The laws governing electric and magnetic fields are summarised most succinctly via the Maxwell equations. The differential forms of which are given below:
\begin{align}
  \nabla \cdot \mathbf{E}
  &= \frac{\rho_ v}{\epsilon_0}
  \\
  \nabla \cdot \mathbf{B}
  &= 0
  \\
  \nabla \times \mathbf{E}
  &= \frac{\partial \mathbf{B}}{\partial t}
  \\
  \nabla \times \mathbf{B}
  &= {\bf j} +\frac{\partial \mathbf{E}}{\partial t}    
\end{align}
where $\rho_v$ is the free charge density, ${\bf j}_0$ is the charge
current, $\epsilon_0$ is the permittivity of free space.

These 4 equations describe how the electric and magnetic fields must behave and are also related to one another, any discussion of optical trapping is underpinned by the fact that in all cases the Maxwell Equations must be satisfied.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Ray-Optics Regime}

The Ray-Optics model is the simplest to understand, the theory models the laser as a collection of individual 'rays' that propagate and are refracted according to Snell's Law. Based on the change in direction momentum is transferred to the target particle; with rays closest to the centre of the beam having greater intensity than those rays at the very edge of the beam. Consider a particle struck by two rays in a Gaussian beam (see Fig below), one coming close to the centre, and the other ray coming from the edge. As each ray is refracted by the target sphere a force is imparted onto it, the total force imparted is given by:
\begin{align}
	F_i = Q_i\frac{\Delta n P_i}{c}
\end{align}
where $Q_i$ is the trapping efficiency, $\Delta n$ is the difference
in refractive indices between the solution and the target particle,
and $P_i$ is the power of the individual ray. For a beam with a
Gaussian intensity distribution $P_i$ will fall off as you move from
the centre of the beam.  The ray optics model is ideal for

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Lornez-Mie Theory}

The Lorenz Mie theory provides an exact solution to the maxwell equations for the scattering caused by an isotropic sphere. The theory describes the scattered wave given off by a dielectric sphere when incident by a plane wave as a summation of partial spherical waves. For any spherical wave the vector fields must solve the Helmholtz wave equation given by:
\begin{align}
	\nabla^2\mathbf{E} +k^2\mathbf{E} = 0
\end{align} 
This combined with the constraints of Maxwell's Equations leaves very few exact solutions apart from spherical or planar waves, both of which can be converted between with relative ease. For example, a plane wave electric field can be expanded into spherical harmonics and likewise any spherical wave can be described as combination of plane waves offset from one another. For a single plane wave the incident, internal, and scattered fields are given as:

@@

\begin{figure}
	\centering
	\includegraphics[width=0.75\linewidth]{./figs/fixed_polarisation.png}
\end{figure}
